Portable Multi- and Many-Core Performance for Finite-Difference/Element Codes; Application to the Free-Surface Component of NEMO (NEMOLite2D 1.0).

### 2017-09-18 docs/papers/nemolite2d_performance/referee_response.txt

Sec 2.1.8 - it is not clear why in-lining manually helps (especially since
one would expect the Intel compiler in particular to do this). We suspect
that the change in the PSy-layer code causes the compiler to make some
different optimisation decisions and these happen to be beneficial. A
proper investigation of this is outside the scope of this paper. We have
added a sentence about this.

Sec 2.2.1 - yes, other strategies might be beneficial, especially blocking.
The key point is that adopting the PSyKAl structure does not limit these
possibilities.

Figure 3 - The Gnu results are much slower because gfortran is only SIMD
vectorising the loop containing the Continuity kernel. For some reason
it does not vectorise any others. Unlike Cray, there is no directive
that can be inserted to mandate that a loop be SIMD vectorised. We have
added text to this effect.

Page 15, line 5 - we were referring to the L3 cache. This has been clarified.

Figure 5 - it is a reflection of the fact that Gnu is free while the Intel and
Cray compilers have performance of the generated code as a key part of their
respective business models.

Page 18, line 9 - the NEMOLite2D performance for the 32x32 domain is
significantly greater than for the other domains because it largely
fits within L2 cache. We have calculated that the working-set size for
the momentum kernel is 136*n^2. For 32x32 this gives ~136KB while for
64x64 this gives ~544KB. The former will fit within L2 cache while the
latter will spill to L3 cache. In addition, recall that we only parallelise
the outermost loop (Sec. 2.2.1) so, for the 32x32 domain, there are only
32 items to be shared amongst the threads. We have added a note to the
text regarding this.

Figure 11 -

### 2017-09-18 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
nemolite2d-paper: respond to comments on serial and OMP performance

Having show that we can recover, and often improve upon, the
performance of the original version of NEMOLite2D, the next logical
step is to examine the necessary code transformations in detail.  We
do this for the $256^{2}$ case since this fits within cache on the Ivy
Bridge CPUs we are using here.  Table~\ref{TABLE_opt_breakdown} shows
detailed performance figures for this case after each transformation
has been applied to the code. The same data is visualised in
Figure~\ref{FIG_opt_stages_256}.

@@

Since the space to explore consists of three different compilers, six
different domain sizes, five stages of optimisation and six different
thread counts we can only consider a slice through it in what follows.
In order to inform our choice of domain size,
Figure~\ref{FIG_omp_domain_size} shows the scaling behaviour of the
most performant, Intel-compiled version of NEMOLite2D. Although it is
common for production runs of NEMO to use MPI sub-domains of $20
\times 20$, it is clear from Figure~\ref{FIG_omp_domain_size} that the
limited quantity of parallelism in the $32 \times 32$ domain inhibits
scaling. Therefore, to fully test the performance of our OpenMP
implementations we have chosen to examine results for the $256 \times
256$ domain and these are shown in Figures~\ref{FIG_gnu_omp_256},
\ref{FIG_intel_omp_256} and \ref{FIG_cray_omp_256}.

>>>>

Having show that we can recover, and often improve upon, the
performance of the original version of NEMOLite2D, the next logical
step is to examine the necessary code transformations in detail.  We
do this for the $256^{2}$ case since this fits within (L3) cache on the Ivy
Bridge CPUs we are using here.  Table~\ref{TABLE_opt_breakdown} shows
detailed performance figures for this case after each transformation
has been applied to the code. The same data is visualised in
Figure~\ref{FIG_opt_stages_256}.

@@

Since the space to explore consists of three different compilers, six
different domain sizes, five stages of optimisation and six different
thread counts we can only consider a slice through it in what follows.
In order to inform our choice of domain size,
Figure~\ref{FIG_omp_domain_size} shows the scaling behaviour of the
most performant, Intel-compiled version of NEMOLite2D. Although it is
common for production runs of NEMO to use MPI sub-domains of $20
\times 20$, it is clear from Figure~\ref{FIG_omp_domain_size} that the
limited quantity of parallelism in the $32 \times 32$ domain inhibits
scaling. (Recall that we are only parallelising the outer loop -
Section~\ref{sec_sep_dos}.) Therefore, to fully test the performance
of our OpenMP implementations we have chosen to examine results for
the $256 \times 256$ domain and these are shown in
Figures~\ref{FIG_gnu_omp_256}, \ref{FIG_intel_omp_256} and
\ref{FIG_cray_omp_256}.
